### 2021-04-30 project_ashish_monograph/CMS.tex
ashish monograph

\section{The CMS experiment}
\label{sec:cms}
(describe the experiment in general, its discovery, mention the international collaboration)

\subsection{CMS detector}
(describe the detector components and what they do)

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./cms12.png}
  \caption{Transverse view of the CMS detector showing the silicon tracker, electromagnetic calorimeter, hadron calorimeter, superconducting solenoid and muon chambers \cite{Chatrchyan:2008aa}.}
  \label{fig:CMS}
\end{figure}

\subsection{The Pixel detector}