### 2021-04-30 project_ashish_monograph/PCC.tex

\section{The Pixel Cluster Counting (PCC) method}
\label{sec:pcc}
(describe cluster reconstruction, the module selection/veto, cluster counting per bx, per LS,..)