### 2021-04-30 project_ashish_monograph/bkg.tex

\section{Backgrounds}
\label{sec:bkg}
(background in vdM, afterglow backgrounds in Physics, subtraction of these)