### 2021-04-30 project_ashish_monograph/vdM.tex

\section{van der Meer calibration}
\label{sec:vdm}