### 2021-05-23 project_ashish_monograph/conclusion.tex

\newpage \section{Summary and Plans} 
\label{sec:conclusion}
sumarize the method and results, and TEPX upgrade work and plans

As plans you can say you will study final module selections, backgrounds, systematics, and sigma_{vis} determination.