### 2022-04-20 tesis_Hedwin/Dedication/dedication.tex

\begin{dedication}
\begin{center}
    \section*{Acknowledgements}
\end{center}
\justify{ 
Thanks to my family for their unconditional support and patience and for always being there when I needed them the most, without them I would not be here.\\
To my friends Carolina, Alejandro and Luisa, for all the help and for making my studies more fun.\\
To my advisor Dr. José Feliciano Benítez,  for always being there to answer my doubts, his patience, and for the opportunity of letting me work in the CMS collaboration. \\
Finally, thanks to CONACYT for supporting my studies with financial aid.}

\end{dedication}