### 2022-04-29 tesis_Rodolfo/Abstract/abstract.tex
Draft_0

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}
    van der Meer calibration is decribed
\end{abstract}