### 2022-04-29 tesis_Rodolfo/Abstract/resumen.tex

\begin{abstract}
    \begin{center}
    {\huge Resumen}
\end{center}
Calibracion de van der Meer
\end{abstract}