### 2022-04-29 tesis_Rodolfo/Chapter4/chapter4.tex

\chapter{Analysis and Results}

\section{Fitted Graphs}
Fit
\section{$\sigma_{vis}$}
xsec