### 2022-12-29 tesis_Rodolfo/Abstract/abstract.tex
Rodolfo Thesis

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}
The calibration of the luminosity measurement from the analysis of the van der Meer scan program by the CMS experiment in the 2018 proton-proton data taking at $\sqrt{s}=$13 TeV with the CMS Silicon Pixel detector is reported. The recorded data was reprocessed using a selection of the Pixel detector corresponding to the period from 2018-05-28 to 2018-07-07, based on luminosity stability studies. The background estimation corresponding to this module selection is reported. The analysis of the van der Meer scans presented here results in the calibration constant $\sigma_{vis}= 9229 \pm 8 (\text{stat.}) \pm 28 (\text{syst.})\text{ mb}$.

\end{abstract}