### 2023-02-16 tesis_Luis/Abstract/resumen.tex

\begin{abstract}
    \begin{center}
    {\huge Resumen}
\end{center}

\end{abstract}