### 2023-02-16 tesis_Luis/Chapter1/chapter1.tex

\chapter{Introduction}
\label{ch1}

%%%%% https://cerncourier.com/a/pushing-the-precision-frontier/
% https://home.cern/news/news/physics/why-precision-luminosity-measurements-matter

\section{Fundamental Particles}

 shown in table \ref{tab:table1}.  
\begin{table}[h!]
  \begin{center}
    \caption{Force experienced by the fermions}
    \label{tab:table1}
    \begin{tabular}{l c c c c}
    &    &  \textbf{Electromagnetic} & \textbf{Weak} & \textbf{Strong}\\
      \midrule[1.1pt]
      \multirow{2}{*}{Leptons} & $e$ \hspace{0.3cm}  $\mu$\hspace{0.3cm} $\tau$ & \checkmark & \checkmark & \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
      & $ \nu_{e} $ \hspace{0.3cm}  $\nu_{\mu}$ \hspace{0.3cm} $\nu_{\tau}$  &  & \checkmark\\ % <-- Content of first column omitted.
      \hline
      \multirow{2}{*}{Quarks} & $u$ \hspace{0.5cm}  $c$\hspace{0.5cm} $t$ & \checkmark & \checkmark& \checkmark\\
      & $d$ \hspace{0.5cm}  $s$\hspace{0.5cm} $b$ & \checkmark & \checkmark & \checkmark\\
%      \hline
%      3&3 & 23.113231 & c\\
%      4&4 & 25.113231 & d\\
    \end{tabular}
  \end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Particle Colliders}

\section{Cross section}

\section{Luminosity }

\section{Importance of Luminosity precision}

\section{The Large Hadron Collider}

\section{LHC Luminosity}

    \label{lumi_per_year_int}
  \end{figure}
\end{center}

%lumi values:
% https://inspirehep.net/files/11b4a1d6569e8e066e7431f17a55d69c

% https://arxiv.org/pdf/1504.06519.pdf

%ATLAS 2015-2018: https://arxiv.org/pdf/1911.04632.pdf