### 2023-02-16 tesis_Luis/Chapter2/chapter2.tex

\chapter{Experiment Description }

\section{The Compact Muon Solenoid}

\section{CMS Tracking System}

\begin{equation}
  -\frac{dE}{dx}= \frac{Dq^{2}n_{e}}{\beta ^{2}} \biggl[ \ln \left(\frac{2m_{e}c^{2}\gamma^{2}}{I} \right) -\beta^{2} -\frac{\delta(\gamma)}{2}  \biggr]
  \label{bethe}
\end{equation}

\subsection{Pixel Detector and Clustering}
\label{pixel_clust_reco}

\subsection{CMS Luminometers}