### 2023-02-16 tesis_Luis/Dedication/dedication.tex

\begin{dedication}
\begin{center}
    \section*{Acknowledgements}
\end{center}
\justify{

\end{dedication}