### 2023-03-01 tesis_Luis/Abstract/abstract.tex
changes

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}
The calibration of the luminosity measurement from the analysis of the van der Meer scanning program for the  CMS experiment in the  2022 proton-proton collision data-taking at $\sqrt{s}=$13.6 TeV with the CMS Silicon Pixel detector is described. The data collection period was on November 11 for the LHC fill 8381, the recorded data collected was reprocessed along with a selection of good modules based on luminosity stability studies. The Analysis of the van der Meer scans with some especific corrections present results in the calibration constant  $\sigma_{vis}= 4163 \pm 3 (\text{stat.}) \pm 12 (\text{syst.})\text{ mb}$.

\end{abstract}