### 2023-03-01 tesis_Luis/Chapter2/chapter2.tex

\chapter{Experiment Description }

\section{The Compact Muon Solenoid}

\section{CMS Tracking System}

\subsection{Pixel Detector and Clustering}

\subsection{CMS Luminometers}