### 2023-03-01 tesis_Luis/Chapter3/chapter3.tex

\chapter{Luminosity  Measurement and Calibration}

\section{Luminosity calibration: van der Meer method}

\subsection*{Backgrounds}