### 2023-03-01 tesis_Luis/Chapter4/chapter4.tex

\chapter{Analysis and Results}

\section{2018 vdM scan program}

\section{Data analysis}

\section{Module selection}

\section{Background estimation}

\section{Corrections}