### 2023-03-01 tesis_Luis/Dedication/dedication.tex

\begin{dedication}
\begin{center}
    \section*{Acknowledgements}
\end{center}
\justify{ 
I would like to thank my wife Elizabeth for her unconditional support in the challenges I set myself, without her this would not be possible. To my advisor, Dr. Jos\'e Ben\'itez, for giving me guidance, patience and, above all, the opportunity to work in such a professional team and join the CMS experiment, that was a dream for me. To the University of Sonora and CONACYT for supporting me in completing this important challenge in my professional career.

\end{dedication}