### 2023-03-29 tesis_Luis/Abstract/abstract.tex
updated

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}
The calibration of the luminosity measurement from the analysis of the van der Meer scanning program for the  CMS experiment in the  2022 proton-proton collision data-taking at $\sqrt{s}=$13.6 TeV with the CMS Silicon Pixel detector is described. The data collection period was on November 10 and 11 for the LHC fill 8381, the data collected was reprocessed along with a selection of good modules based on stability studies. The analysis of the van der Meer scans with some especific corrections results in the calibration constant  $\sigma_{vis}= 4163 \pm 3 (\text{stat.}) \pm 12 (\text{syst.})\text{ mb}$ for the Pixel Cluster Counting luminometer.

\end{abstract}