### 2023-06-16 tesis_Ashish/Abstract/abstract.tex
Your commit message

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement during LHC run periods. The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker. PCC is studied for its stability and accuracy in luminosity measurements using 2018 AlCa Reco datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and accurate, able to achieve ??\% systematic uncertainty on the final luminosity measurement for Run 2 data. Moreover, the algorithms and optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using the Tracker Endcap Pixel Detector (TEPX). This detector is expected to provide excellent statistical and linearity performance, which is essential for accurate luminosity measurements at the HL-LHC. The TEPX detector will replace the upgraded Phase I pixel detector and is expected to provide even higher precision luminosity measurement.

\end{abstract}