### 2023-06-16 tesis_Ashish/Acknowledgement/acknowledgement.tex

I would like to express my sincere gratitude to all those who have supported and guided me throughout my journey in completing my PhD thesis in particle physics, specifically focusing on the CMS experiment at CERN. Their unwavering support, encouragement, and expertise have been invaluable to my success.

First and foremost, I would like to extend my deepest appreciation to my esteemed thesis supervisor, Dr. Jose Feliciano Benitez. His exceptional guidance, profound knowledge, and unwavering commitment to scientific exploration have shaped my research endeavors and allowed me to delve into the fascinating world of particle physics. His mentorship and continuous support have been instrumental in every step of my thesis, from conceptualization to the final stages of completion. I am truly grateful for the opportunity to work under his supervision, and I have grown immensely both as a researcher and as an individual thanks to his invaluable guidance.

I would also like to express my heartfelt gratitude to my loving parents, whose unconditional love, encouragement, and sacrifices have been the cornerstone of my academic journey. Their unwavering belief in my abilities and their constant support have given me the strength and motivation to overcome challenges and pursue my passion for scientific exploration. I am eternally grateful for their constant presence in my life and their unwavering support throughout this demanding academic endeavor.

Furthermore, I would like to extend my appreciation to my sisters for their unwavering encouragement, understanding, and motivation during the course of my PhD. Their belief in my abilities and their continuous support have been an immense source of inspiration and strength, reminding me to remain focused and determined even during the most challenging times.

I would also like to acknowledge the collective efforts of the scientific community at CERN and all those who have contributed to the CMS experiment. Their collaboration, expertise, and dedication have created an enriching environment for research and have allowed me to be a part of groundbreaking discoveries in the field of particle physics.

Lastly, I am deeply grateful to all my friends and colleagues who have provided valuable insights, stimulating discussions, and endless support throughout this journey. Their encouragement, camaraderie, and intellectual exchanges have played a significant role in shaping my research and personal growth.

To each and every person mentioned above, as well as those who may not be named individually, I offer my sincerest thanks. Your contributions have been invaluable, and I am humbled by the generosity and unwavering support I have received. This thesis would not have been possible without your guidance, encouragement, and belief in me.

Thank you from the bottom of my heart.