### 2023-06-16 tesis_Ashish/Chapter8/chapter8.tex

\section{Summary and Plans} 
\label{sec:con}

\subsection{Summary of the study and its findings}
\subsection{Limitations and potential improvements of the pixel cluster counting method}
\subsection{Future work related to luminosity measurement at the LHC}
\subsection{Conclusion and final remarks}