### 2023-07-05 tesis_Ashish/Abstract/abstract.tex
AS thesis update

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}

The Pixel Cluster Counting (PCC) luminosity measurement method has emerged as a reliable offline technique for luminosity measurement.
%The PCC method is based on the rate of pixel clusters in the CMS pixel detector, which is the innermost part of the CMS tracker.
PCC method is studied for its precision in luminosity measurement using 2018 AlCa Reco datasets. The results of these studies demonstrate that the PCC luminosity measurement method is stable and linear, able to achieve 1\% systematic uncertainty on the final luminosity measurement. Moreover, the clusters and coincidences counting algorithms and their optimization for the Phase II High Luminosity LHC (HL-LHC) luminosity measurement are also developed, using simulated datasets for the proposed Phase II pixel detector. %This new pixel detector will replace the Run 2 pixel detector.
Linearity of the TEPX and D4R1 luminometers are expected to be within 1\% (offline) up to a pileup of 200 and their statistical precision per second is about 0.1\% at <PU>=200.
%It is expected to provide excellent statistical precision and linearity performance, which is essential for precision luminosity measurements at the HL-LHC.

\end{abstract}