### 2023-12-17 tesis_Ashish/Chapter8/chapter8.tex

\chapter{Future work}  %Title of the First Chapter

\ifpdf
    \graphicspath{{Chapter8/Figs/Raster/}{Chapter8/Figs/PDF/}{Chapter8/Figs/}}
\else
    \graphicspath{{Chapter8/Figs/Vector/}{Chapter8/Figs/}}
\fi

\section{Summary and Plans} 
\label{sec:con}

\subsection{Summary of the study and its findings}
\subsection{Limitations and potential improvements of the pixel cluster counting method}
\subsection{Future work related to luminosity measurement at the LHC}
\subsection{Conclusion and final remarks}