### 2024-03-27 tesis_Ashish/Declaration/declaration.tex
update 28 Mar

\begin{dedication}
  \vspace*{-2cm} % Adjust the -2cm to the amount of space you want to move up                                                                                                                                                                
%\centering                                                                                                                                                                                                                                  
%\begin{center}                                                                                                                                                                                                                              
\section*{Declaration}
%\end{center}                                                                                                                                                                                                                                
\justify{

I, Ashish Sehrawat, solemnly affirm that the doctoral thesis entitled "Measurement of the Luminosity in Proton-Proton Collisions at the CMS Experiment of the Large Hadron Collider," submitted for the completion of the Doctorado en Ciencias (Física) at the Universidad de Sonora, is my original work. It has not been submitted for any degree or examination at another university.

The results and discussions presented in Chapter 5 of this thesis are part of the collaborative work published as “Luminosity measurement in proton-proton collisions at 13.6 TeV in 2022 at CMS,” CMS Collaboration, 2024, CMS PAS LUM-22-001, accessible at https://cds.cern.ch/record/2890833.

The outcomes of Chapter 6 are included in “The Phase-2 Upgrade of the CMS Beam Radiation Instrumentation and Luminosity Detectors,” by the CMS Collaboration, technical report, CERN, Geneva, 2021, available online at https://cds.cern.ch/record/2759074. These contributions are properly cited and recognized within this document.

This thesis has been prepared with strict adherence to ethical standards of academic conduct. All sourced information has been appropriately cited to respect the intellectual property rights of original authors. This work has been supported by the Universidad de Sonora and the Consejo Nacional de Ciencia y Tecnología (CONACYT), to whom I extend my deepest gratitude for their support.

By signing below, I certify the authenticity and originality of the work submitted herein and affirm all statements made as true and based on my research findings.

Signature:

Ashish Sehrawat

Date: 28 March 2024

Location: New Delhi, India

\end{dedication}