### 2025-03-07 monograph_Luis/Abstract/abstract.tex
monografia

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}

\end{abstract}