### 2025-03-07 monograph_Luis/Chapter1/chapter1.tex

\chapter{Introduction}
\label{ch1}

\section{Particle Physics and The Standard Model}

\section{Particle Colliders and production cross section}

\section{Luminosity }

\section{The Large Hadron Collider}

\section{Expected Luminosity precision for LHC Run 3}