### 2025-03-07 monograph_Luis/Chapter2/chapter2.tex

\chapter{Experiment Description}

\section{The Compact Muon Solenoid}

\section{CMS Luminometers}
\label{Luminometers}

\section{CMS Tracking System}

\section{Silicon Pixel Detector}

\section{Pixel Cluster Reconstruction}