### 2025-03-07 monograph_Luis/Chapter3/chapter3.tex

\chapter{Luminosity  Measurement and Calibration}
\label{ch3}

\section{2017 and 2018 CMS Data collected}

\section{Module selection}

\section{Luminosity calibration: van der Meer method}