### 2025-03-07 monograph_Luis/Chapter4/chapter4.tex

\chapter{Analysis and Results}

\section{2017 and 2018 vdM scan program}
\label{2017 and 2018 vdM scan program}

\section{Data acquisition and processing}
\label{data}

\section{Background estimation}
\label{bkg}

\section{Beam Corrections}
\label{bbcorrections}
%https://cms.cern.ch/iCMS/analysisadmin/cadilines?line=LUM-22-001

\section{Fit Model  Selection}

\section{Results}

\section{Systematic Uncertainties}