### 2025-04-21 thesis_julio/Chapter4/chapter4.tex

\chapter{Results}
\label{ch4}

\section{Sigma_Vis}
\section{Systematics}
\sectin{}