### 2025-05-06 monograph_Luis/Abstract/abstract.tex

\begin{abstract}
\begin{center}
    {\huge Abstract}
\end{center}

This work presents the calibration of the CMS Pixel Cluster Counting (PCC) luminometer and the absolute luminosity determination using the van der Meer (vdM) method, as part of the publication \textit{``Precision luminosity measurement in proton-proton collisions at $\sqrt{s} = 13$ TeV with the CMS detector''}, which covers the second part of Run 2 corresponding to the 2017 and 2018 data-taking periods. The final results for the PCC visible cross section $\sigma_{\text{vis}}$ are 4666.4 $\pm$ 14.7 mb for 2017 and 964.9 $\pm$ 2.2 mb for 2018. The final integrated luminosity is obtained from a combination of multiple independently calibrated luminometers. A precision of 0.86\% for 2017 and 0.83\% for 2018 is achieved, with dominant uncertainties arising from luminometer non-linearity and the x–y factorization correction.

\end{abstract}