### 2025-06-11 thesis_julio/Chapter3/chapter3.tex
10june

\chapter{Luminometer Calibration}
\label{ch3}

\section{Van Der Meer Method}

The Van Der Meer (vdM) scan let us obtain the value of $\sigma_{visible}$ using different luminometers. For this case the detector of interest is the pixel detector with which are going to obtain a calibration of the luminosity using the vdM scan in combination with the PCC method. The scan consist on separating particle the beams from each other between the X and Y axis by $\Delta X$ and $\Delta Y$ values moving them and recording values of luminosity independently, when the scan on X and scan on Y are imposed over each other we said that they're on head on giving the maximum value of luminosity. Usually a single gaussian fit is used to obtain beam overlap widths also denoted as $\Sigma_{x}$ and $\Sigma_{y}$, other fit models may be used like the doble gaussian depending on the quality of the fit \cite{Vdm}

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{vdm1.png}
    \caption{The vdm scan on the left showing the beams for X and Y. On the right a curve of the different values of rate for the beams displacement, the peak occurs during the head on part.}
    \label{fig:vdm1}
\end{figure}

For obtaining the importnat parameters on the vdm scan we return to the equation (1.2) and apply it to and individual bunch crossing, the values of N1 and N2 being the number of protons colliding can be obtained since is a known quantity from the experiment, the frequency f is also a known quantity, the frequency of the LHC which is 11,245 kHz but the proton densities $\rho_{1}$ and $\rho{2}$ are more difficult to obtain, this is were the vdM scan help us to measure integral over the bunch proton densities. The luminosity integral evaluated with the beams separated by a distance $\Delta_{x}$ and $\Delta_{y}$ can take the form of:

\begin{equation}
 L(\Delta x, \Delta Y) = N_{1} N_{2} f  \int \int \rho_{1}(x,y)\rho_{2}(x+\Delta x, y+\Delta y) dxdy 
\end{equation}

Because of the scan method is it assumed that the two bunches of proton densities are factorizable since they act independently between each other so we can turn the right part of (3.1) into two integrals:

 \begin{equation}
N_{1} N_{2} f  \int \int \rho_{1}(x,y)\rho_{2}(x+\Delta x, y+\Delta y) dxdy   = N_{1} N_{2} f (\int \rho_{1}(x)\rho_{2}(x + \Delta x) dx) (\int \rho_{1}(y) \rho_{2}(y + \Delta y) dy)
\end{equation}

Integrating both sides on $\Delta y$ and using in combination with (3.1) while we fix $\Delta x_{0}$ = 0 and $\Delta y_{0}$ = 0 as the head on point,  we obtain:

 \begin{equation}
N_{1} N_{2} f \int \rho_{1}(x) \rho_{2}(x + \Delta x_{0}) dx = \int L (\Delta x_{0}, d\Delta y) d(\Delta y)
\end{equation}

A similar result is obtained when we integrate both sides for $\Delta x$ using this and combining equation (3.1), (3.2) and (3.3) we can obtain that

\begin{equation}
N_{1} N_{2} f \int \rho_{1}(x) \rho_{2}(x + \Delta x_{0}) dx = \frac{L (\Delta X_{0}, \Delta y_{0})}{\int L(\Delta x_{0}, \Delta y) d(\Delta y)}
\end{equation}

Similarly to the previous step we can obtain the value of the other factor of (3.2) with an analogous process. The integrals resulting on the right side can be evaluated by evaluated by obtaining the rate in function of the beam-beam separation which are $\Delta x$ and $\Delta y$  this because the luminosity has a linear relationship with the rate. The Luminosity can be expresed on terms of the rate during the head on points using the following:

\begin{equation}
L(\Delta x, \Delta y) = N_{1} N_{2} f \frac{2R(\Delta x_{0}, \Delta y_{0}}){\int L(\Delta x_{0}, \Delta y) d(\Delta y) \int L(\Delta x, \Delta y_{0}) d(\Delta x)}
\end{equation}

Here the luminosity is replaced by the rate R, it is convenient to write this integral in terms of the convoluted beam widths which are denoted by $\Sigma_{x}$ and $\Sigma_{Y}$, this convoluted widths are defined by:

\begin{equation}
\Sigma_{x} = \frac{1}{2 \pi} \frac{\int R(\Delta x, \Delta y_{0} d(\Delta x) )}{R(\Delta x_{0}, \Delta y_{0})}
\end{equation}

For $\Sigma_{y}$ the result is analogous. With (3.6) we can write (3.5) in a different manner:

\begin{equation}
L(\Delta x, \Delta y) = \frac{N_{1}N_{2}f}{2 \pi \Sigma_{x} \Sigma_{y}}
\end{equation}

This in combination with (1.4) make us possible to obtain the following expression for the visible cross section

\begin{equation}
\sigma_{vis} = \frac{2 \pi \Sigma_{x} \Sigma_{y} R(\Delta x_{0} \Delta y_{0})}{N_{1} N_{2}f}
\end{equation}

This makes possible obtain $\sigma_{vis}$ only by experiment since all of the right values can be obtained via the experiment, the maximum rate R($\Delta x_{0}$, $\Delta y_{0}$) is obtained at the maximum PCC obtained during the vdm scan, this means during the head-on process, while the $\Sigma_{x}$ and $\Sigma_{y}$ are obtained with the information of the fit model.

This calculations are done on several Bunch Cross Identifier which so you obtain several values of $\simga_{vis}$ and then they are weight averaged according to the uncertainties in order to obtain $\sigma_{vis}$ for the scan.

\section{Datasets}

The data on the CMS is a complex set of inter-dependent workflows made in a way that assure the full physics exploitation of the CMS detector potential and the collisions delivered by the LHC, this data provides analyses with reconstructed collision events from the experiment and are designed to use the CMS computing resources efficiently. This stream of events is organized into datasets according to the results of the High Level Trigger (HLT) which defines primary datasets since they are defined by the paths of the HLT. The design of the primary datasets is centered around candidates for particles that are reconstructed on the final state by the hLT and follows the principle of grouping together events with similar physics content.    \cite{datasets1}

The datasets used for the obtention of the $\sigma_{visible}$ for 2024 were the ones from the Fill 9639 which is divided in 3 blocks, a fill is a process that generates beams for the LHC this tipycally involves particles around the number of $10^{14}$ that are grouped in bunches that form the proton beam. This is what we use the bcid number to identify this bunches, there are a total of 3564 bunches that are empty or filled depending on the fill all of this part of the LHC bunch train. For this fill in the calibration the bcids of relevance were a total of 12 which are identified by the numbers: 303, 324, 345, 506, 527, 822, 1081, 1102, 1397, 2000, 2965, 3123.

In the case for our datasets each bcid is processed by the CMS collaboration to enable cluster reconstruction resulting in the ALCARECO datasets that contain a module collection and their number of cluster per event in CMSWW format. This samples are processed using the CMSWW software to extract the clusters per module to obtain the dataset in the ROOT format. ROOT is a software framework born at CERN It is open source and used for high energy physics to analyze data while providing packages for storage, processing and visualization while minimizes the computing resources needed.

The root format saves the data in form of TTrees which behaves like an array of data structure and storage data. This tree consist on branches and leaves, a branch is a list of independent columns that can contain values of any fundamental type and are represented by TBranch. While the leaves represented by TLeaf give access to actual data in difference to the TBranch which represent a structure. After the datasets are processed, the rates of the Pixel detector were stored each 1.32 seconds periods called NB4 this to make an average of clusters and the number of events that fall into that time span.

This dataset were first processed by the CMS collaboration enabling a cluster reconstruction resulting in the ALCARECO datasets getting a lot of root files for each of the different zero bias streams that were the starting point of this analysis.

After this the data then was processed again to obtain a hd5file a format that can supports n-dimensional datasets and facilitates the reading of the datasets, this hd5 files contained relevant information about the root files like the average rate per NB4, this was made for each of the files for almost all of the available streams excluding two of them that were not ready by the moment (the ones that were identified by ZB21 and ZB28) after one stream got all of his files processed into an hd5 file they were converged into a single file containing relevant information for all the stream, once all of the files in all of the data streams were processed into hd5 files a single hd5file was made containing all of the information about the streams also a single csv file was made containing only the average rate file per NB4 for each of the bcids.

\section{Scans}

For the calibration of 2024 in the LHC fill 9639 during may 15 to may 16 at collision proton proton with a energy of 13.6 TeV, the vdm scans were made there were a total of 5 vdm X and vdm Y scans and 2 beam img X and beam img Y and the Super Separation period which is a scan used to obtain a background estimation, this were the scans used on this analysis to compute the value of $\singa_{visible}$. The figure (3.x) shows a DOROS image of the fill and all the scans that were made during that period of time.

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth, height=0.4\textheight]{scans.png}
    \caption{Fill 9639 scans}
    \label{fig:scans}
\end{figure}

The 5 vdm scans vdM1, vdM2..., vdM5 are the scans usually used in the calibration to compute $\simga_{visible}$ two beams are separated the transverse bunch size and then scanned across one another in sequence of 25 steps with 30 seconds per step. The beam imaging scan is kept on a nominal position while the other is separated and scanned in 19 steps, the scan is performed first in the X direction and then in the Y direction. The Super Separation scans separates the beams a distance were the overlap of the beams is negligible, the minimum possible.

\section{Background Estimation}

The background estimation were obtained using the 5 Super Separation scans, the time on which this super separations were used was on UNIX time each of these super separations lasted a total of 5 minutes in the following table we got more information about the start and ending of the scans.

\begin{table} [H]
\begin{center}
\caption{Super Separation Period time}
\begin{tabular}{|c c c|} 
 \hline
 Scan & Time Start & Time End  \\ [0.5ex] 
 \hline\hline
 SS1 & 1715883182 & 1715883481  \\ 
 \hline
 SS2 & 1715902185 & 1715902484  \\
 \hline
 SS3 & 1715919164 & 1715919463 \\
 \hline
 SS6 & 1715943045 & 1715943343  \\
 \hline
 SS5 & 1715987303 & 1715987602  \\ [1.0ex]
 \hline
\end{tabular}
\end{center}
\end{table}

The data of this scans were processed and analyzed, in the following image we show the mean value of the PCC rates during the Super Separation period scan in one bcid.

\begin{figure}[H]
    \centering
    \includegraphics[width=1\textwidth]{SS1.jpeg}
    \caption{Mean rates during the second super separation scan}
    \label{fig:SS1}
\end{figure}

The reason why the background estimation is obtained during the super separation (SS) process, It's because the SS scan separates the beams so they shouldn't be colliding the expected value on the rate is expected to be a rate associated with the background, this was obtained checking the mean values for each bcid for all the entries of pcc rates for each of the 5 periods of super separation, the following image illustrate a case for one SS period and bcid 50x:

\begin{figure}[H]
    \centering
    \includegraphics[height=0.4\textheight, width=0.8\textwidth]{sshist.png}
    \caption{SS periods histograms showing entries for bcid 345}
    \label{fig:SShist}
\end{figure}

The mean value during this period is assumed to be the background noise value of this SS period, in the following table we see the mean PCC value for all of the bcids and how it changes in each of the five super separation periods.

\begin{figure}[H]
    \centering
    \includegraphics[height=0.30\textheight,width=1\textwidth]{SSmean.png}
    \caption{Mean per bcid}
    \label{fig:SSmean}
\end{figure}

In principle this value is a single value so we obtain a histogram obtaining a mean for all the mean values for all the 60 values this is shown in the following image:

\begin{figure}[H]
    \centering
    \includegraphics[width=1\textwidth]{ssthist.png}
    \caption{Histogram of all means}
    \label{fig:Histogram}
\end{figure}

With this que obtained that the value for super separation is $ss_{background}$ = 0.0251 +-0.0018 this value would be extracted from the PCC rates during the fit for the vdm scans.

\section{Corrections}

A few corrections were taken into consideration to eliminate noise data from another sources during this vdm calibration, this corrections are applied by the vdm framework.

\begin{itemize}
  \item \textbf{Background} The rates substracted on the SS periods and is associated with the luminometer itself see section (3.x) for more.
  \item \textbf {Ghost Satellite} This correction have to be with the charge that comes from the non colliding bunches, which are called ghost and those who are leaking from the main radio frequency bucket into another bucket which are called satellite. This currents have an effect on the $\sigma_{visible}$ measure because of the contribution on the charges, this charges are substracted from the beam currents.  \cite{ghost}
  \item \textbf{BeamBeam} This are deflections that happen between bunch crossings, accounting for the electrical repulsion of the beams which increases the lateral separation.   \cite{beambeam}
  \item \textbf{Dynamic beta} This correction accounts for any changes in the proton density distributions of the bunches due to the single particle interaction, resulting into a non linear change during separation steps in transverse bunch profiles and can be described as the effective changes of a called $\beta$ value. \cite{LHClum}
\end{itemize}

\section{Fit model}

There are different fit models at the time to do the vdm calibration, the most common one is Single Gaussian but also other models like the Double Gaussian, QG, Poly2g or Poly 4G. Each of these models have different parameters and restrictions making some of them better than other on the fit quality.

For the main calibration we used the QG fit this fit uses the parameter $\sigma$, the peak and a q values as the parameters. This fit is given by the following equation:

\begin{equation}
f_{qg}(x,y) = peak e_{q}(-\beta(x - mean)^{2},q)
\end{equation}

with $e_{q}$ being a function of x defined as:

\begin{equation}
e_{q}(x) = (1+(1-q)x)^{\frac{1}{1-q}}
\end{equation}

q being a parameter and peak the maximum rate.

This fit model were applied to the data using vdmfamework software for all bcids, the following image shows this fit on the x plane and on the y plane on one bcid other bcids showed a similar result.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{fitqg1.png}
    \includegraphics[width=0.45\textwidth]{fitqg2.png}
    \caption{QG Fit on X plane and Y plane}
    \label{fig:QGfit}
\end{figure}

To see if the fit converged we check the covariance matrix with the covalue, if this value is 3 the fit converge as we can see in the following image all of the bcids converged in all of the scans for the vdm and BI. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{QG_covStatus.png}
    \caption{Covariance value showing that all bcids converged with the fit}
    \label{fig:QGfit}
\end{figure}

Also to see the fit quality we obtain the value of $\chi^{2}$ divided by the numbers of degree freedom the following image shows this:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{QG_chi2.png}
    \caption{Fit quality values}
    \label{fig:QGchi2}
\end{figure}

\section{Beam Parameters}

The beam parameters of the vdm scan were obtained after the vdm framework run on the fit, one of the parameters obtained is the peak:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{QG_peak.png}
    \caption{QG peak values important on the computing of $\sigma_{visible}$}
    \label{fig:QGpeak}
\end{figure}

Here we can see that the peak values are increasing with the time which doesn't have sense since the number of particles through time should be lower, this behavior is then explained if we observe the results for the $\Sigma_{x, y}$ in the following image:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{QG_CapSigma.png}
    \caption{QG $\Sigma$ important to compute value of $\sigma_{visible}$}
    \label{fig:QGcapsigma}
\end{figure}

Here we see that the $\Sigma_{x, y}$ values are decreasing for each scan reducing the thickness of the beams and thus observing a higher peak value even if the number of particles is decreasing through time.

Now with all the values obtained we can compute the value of $sigma_{visible}$, the values for each bcid is shown on the following image:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{sigbcid.png}
    \caption{QG $\Sigma$ important to compute value of $\sigma_{visible}$}
    \label{fig:sigbcid}
\end{figure}

We are expecting a constant value of $\sigma_v{visible}$ here it appears to be changing each scan this could be to other uncertainties. Then in the following image we show the result for the weighted mean of the $\sigma_{visible}$ for each scan:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{sigscan.png}
    \caption{QG $\Sigma$ important to compute value of $\sigma_{visible}$}
    \label{fig:sigscan}
\end{figure}

Which give us a $\sigma_{visible}$ value of

\begin{equation}
\sigma_{visible} = 1460890 \pm 546
\end{equation}

Which is the value for $\sigma_{vis}$ that we have been looking for the calibration using the vdm scans.

\newpage

\section{Systematics}

We obtained a systematic uncertainties comparing the fit made but with different corrections, in the following table we check this uncertanties with citing a relative change per correction:

\begin{table}[H]
\begin{center}
\small
\begin{tabular}{|c c c|} 
 \hline
 Correction & $\sigma_{visible}$ & Relative Change(\%) \\ [0.5ex] 
 \hline\hline
 No correction & $1452589 \pm 2041$ & - \\ 
 \hline
 Background & $1448801 \pm 676$ & \\
 \hline
 Background + Ghost Satellite & $1450154 \pm 560$ & \\
 \hline
 Background + Ghost Satellite + Beam Beam + Dynamic Beta & $1460890 \pm 546$ & \\ [1.0ex]
 \hline
\end{tabular}
\caption{Comparison of $\sigma_{visible}$ values with different corrections}
\end{center}
\end{table}