### 2025-06-24 thesis_julio/Abstract/abstract.tex
thesis24june

\begin{abstract}
\begin{center}
     {\huge Abstract}
\end{center}

This job reports the calibration of the luminosity measurement using the analysis of the van der Meer scan program by the CMS experiment in 2024 proton-proton collision at $\sqrt{s}$ = 13 TeV with the Silicon Pixel detector in the CMS. The data used and processed for this analysis was during the period of May 16 to May 17 corresponding to the LHC fill 9369. The background was obtained and the analysis on the van der Meer scan with some corrections results in the calibration constant $\sigma_{vis}$ = 1460890  \pm 546 (stat) \pm 7304(syst)$ $\mu b$.

\end{abstract}